---
title: "chapter2metaanalysis"
author: "Nicole Peckham"
date: '2022-12-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(janitor)
library(here)
library(tidyverse)
library(readxl)
library(ggplot2)
library(sciplot)
library(reshape2)
library(lubridate)
library(broom)
library(ggpubr)
library(lme4)
library(car)
library(report)
library(forcats)
library(lmerTest)
library(MuMIn)
#library(waffle)
library(ggeffects)
library(patchwork)


overall_p <- function(my_model) {
    f <- summary(my_model)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}


#setwd("~/local R files/Ch2/metaanlysis")

```

```{r meta analysis figures, include=FALSE}

all <-  read_xlsx("tempmaster6aug2026.xlsx", sheet = "master_rows")
# import metaanalysis data for temperature

blank<-read.csv("blank.csv")
blank <- blank %>%
  mutate(lrr = es)

colnames(all) <- as.character(all[1,])
all <- all[-1,]
all <- clean_names(all)
all$year <- as.numeric(all$year)

all$risk_neg_error_val <- as.numeric(all$risk_neg_error_val)
all$risk_val <- as.numeric(all$risk_val)
all$risk_pos_error_val <- as.numeric(all$risk_pos_error_val)
all$risk_n <- as.numeric(all$risk_n)

all$no_risk_neg_error_val <- as.numeric(all$no_risk_neg_error_val)
all$no_risk_val <- as.numeric(all$no_risk_val)
all$no_risk_pos_error_val <- as.numeric(all$no_risk_pos_error_val)
all$no_risk_n <- as.numeric(all$no_risk_n)

all$no_context <- as.factor(all$no_context)

all$num <- as.factor(all$num)
all$iso <- as.numeric(all$iso)



all <- all %>%
  mutate(response_type = str_replace(response_type, "reproduction", "population growth"))

adj <- read.csv("TPC.csv")

adj <- adj %>%
  select(id, adjust) %>%
  unique()

all <- all %>%
  left_join(adj, by = "id") %>%
  mutate(adjust = replace_na(adjust, 0))%>%
  mutate(temp = iso+adjust)


```



```{r}

paperstable_old <- all %>%
  mutate(id = as.factor(id)) %>%
  dplyr::group_by(id) %>%
   dplyr::select("id", "year", "authors","complete", "strict", "nce_tmie", "scale", "no_context", 
                 "response_type",
                 "prey_taxa", "pred_spec","prey_spec", "system", "temp") %>%
  filter(strict == "y")

range <- paperstable_old %>%
  dplyr::group_by(id) %>%
  dplyr::summarize(min = min(temp), max = max(temp)) %>%
  mutate(range = max - min)

paperstable <- paperstable_old %>%
  filter(complete == "Y") %>%
  dplyr::select(-temp) %>%
  unique() %>%
  merge(range)


write.csv(paperstable, file="paperstable.csv")



marine_multiple <- paperstable_old %>%
  filter(system == "m") %>%
  dplyr::select(-temp) %>%
  unique()

```



```{r confidence intervals}
# using manually calculated errors from spreadsheet


all$se_val_risk <- as.numeric(all$se_val_risk)
all$se_val_no_risk <- as.numeric(all$se_val_no_risk)


# need to shift two studies because can't LRR w/ 0 or negative

# 1) shift the vals 

shift <- 1
zero <- all %>%
  filter(risk_val <= 0) %>%
  mutate(risk_val = risk_val+shift, 
         risk_neg_error_val= risk_neg_error_val+shift,
         risk_pos_error_val= risk_pos_error_val+shift,
         no_risk_val= no_risk_val+shift, 
         no_risk_pos_error_val = no_risk_pos_error_val+shift,
         no_risk_neg_error_val = no_risk_neg_error_val+shift)

# 2) remove original from paper and use this new one instead

all <- all %>%
  filter(risk_val > 0) %>%
  bind_rows(zero)

# done!

w_error <- all %>%
  filter(complete == "Y") %>%
  dplyr::select(num, id, year, authors, nce_tmie, scale, strict, no_context, system, prey_taxa, 
                 response_type, specific_response, context_type, c_treatment, temp, risk_val, risk_error_type,
                no_risk_val, se_val_risk, se_val_no_risk, risk_n, no_risk_n) %>%
  mutate(lrr = log(risk_val/no_risk_val)) %>%   #LRR
  mutate(v = ((se_val_risk*sqrt(risk_n))/(risk_n*risk_val)) + 
           ((se_val_no_risk*sqrt(no_risk_n))/(no_risk_n*no_risk_val))) %>%
  filter(strict == "y")  # filter to only strict interpretation papers 




w_error$prey_taxa <- as.factor(w_error$prey_taxa)

 
```


# fix labels

```{r}

w_error$system <- fct_recode(w_error$system, "Freshwater" = "fw", "Marine" = "m", "Terrestrial" = "t")
w_error$scale <- fct_recode(w_error$scale, "Field" = "field", "Lab" = "lab")


paperstable$system <- fct_recode(paperstable$system, "Freshwater" = "fw", "Marine" = "m", "Terrestrial" = "t")
paperstable$scale <- fct_recode(paperstable$scale, "Field" = "field", "Lab" = "lab")
paperstable$response_type <- fct_recode(paperstable$response_type, "Density" = "density", 
                                        "Fecundity" = "fecundity", "Growth" = "growth", 
                                        "Mortality" = "mortality", 
                                        "Population Growth" = "population growth")
paperstable$system <- as.factor(paperstable$system)

```

# groupings

```{r}

papers <- paperstable %>%
  dplyr::select(authors, nce_tmie, system, scale,  min, max, range) %>%
  unique()
tot.papers <- nrow(papers)
tot.expts <- nrow(paperstable)

nrow(paperstable[paperstable$nce_tmie == "NCE", ])
nrow(paperstable[paperstable$nce_tmie == "NCE", ])/tot.expts

nrow(paperstable[paperstable$nce_tmie == "TMIE", ])
nrow(paperstable[paperstable$nce_tmie == "TMIE", ])/tot.expts


# how many of each system?
paperstable %>%
  dplyr::group_by(system) %>%
  count() %>%
  mutate(perc = n/tot.expts)

# how many of each response type?
paperstable %>%
  dplyr::group_by(response_type) %>%
  count() %>%
  mutate(perc = n/tot.expts)

# how many in response type/system?
paperstable %>%
  dplyr::group_by(system, response_type) %>%
  count() %>%
  mutate(perc = n/tot.expts)


## how many publications did we find?

npubs <- paperstable %>%
  dplyr::select(id) %>%
  unique()

# how many of nce/tmie?
paperstable %>%
  dplyr::group_by(nce_tmie) %>%
  count() %>%
  mutate(perc = n/tot.expts)


```

# ALL STUDIES

```{r}
# Figure 1. 

paperstable$response_type=factor(paperstable$response_type, levels=c("Growth", "Mortality",
                                                                     "Fecundity", 
                                                                     "Population Growth",
                                                                     "Density"))
                                      
```

## setup figure

```{r}


example <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Fitness") +
    xlim(-5,5)+    
  coord_cartesian(ylim = c(-3,3)) +  
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
example

ggsave(plot=example, filename = "example.png", width=4, height =4)


```

## - figure 1

```{r}

str(paperstable)

fig1 <- 
paperstable  %>%
  ggplot() +
  geom_bar(mapping=aes(x=response_type,group=nce_tmie,fill=nce_tmie), color= "black") +
  facet_wrap(~system) +
   coord_equal() +
  theme_bw() +
  scale_fill_manual(name= "Scale", values = c("gray", "black")) +
  scale_y_continuous( expand = c(0, 0),  breaks = c(0, 2, 4, 6, 8, 10)) +
  xlab("Response Type") +
  ylab("Number of Studies")+
  ylim(c(0,11))+
    theme(panel.background = element_rect(fill = "white"),
          strip.text.y.left = element_text(),
          axis.text.x=element_text(angle = 45, hjust = 0.9, vjust = 1),
    panel.grid.minor = element_blank()
    )
fig1

ggsave(plot=fig1,device = 'pdf',filename = "fig1.pdf", width=7, height = 5)


```

```{r}

w_error %>%
  select(system, temp) %>%
  distinct %>%
  group_by(system) %>%
  summarize(xmin=min(temp), xmax=max(temp))

w_error %>%
  dplyr::select(system, lrr) %>%
  distinct() %>%
  group_by(system) %>%
  summarize(ymin=min(lrr), ymax=max(lrr))


ranges <- w_error %>%
  group_by(nce_tmie, response_type, system) %>%
 dplyr::summarize(min_temp = min(temp, na.rm = TRUE),
                   max_temp = max(temp, na.rm = TRUE)) %>%
  mutate(range = abs(min_temp) + abs(max_temp))



```







Mixed models were fit with the lme4 package (Bates
et al. 2015). Furthermore, in models with significant main effects
of site, predator cue, and/or CE, we calculated marginal
R2 values for each effect in order to evaluate the relative
influence of each factor on the response variable
(MuMIn package; Barton 2019).

There are two types of \RsqxGLMM: marginal and conditional.

Marginal \RsqxGLMM represents the variance explained by the fixed effects, and is defined as:

Conditional \RsqxGLMM represents the variance explained by the entire model, including both fixed and random effects. It is calculated by the equation:


# growth

## - Figure 2
```{r}


rawgrowth <- w_error %>%
  filter(response_type == "growth")


growth_mod <- lmer(lrr~temp + (1 | id), data=rawgrowth)
summary(growth_mod)
Anova(growth_mod)

r.squaredGLMM(growth_mod)
shapiro.test(lm(lrr~temp,data = rawgrowth)$residuals)

# color palette
N=length(unique(rawgrowth$id))
Cols = colorRampPalette(brewer.pal(8, "Accent"))

figgrowth <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=rawgrowth, mapping=aes(x=temp, y=lrr, color=id), 
             position=position_dodge(width=0.8)) +
    scale_color_manual(values=Cols(N))+

  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
#    coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(data=rawgrowth, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-sqrt(v), ymax = lrr+sqrt(v)), 
                position=position_dodge(width=0.8)) +
  theme_bw() +
  # geom_text(data = rawgrowth, aes(x=10.5, y= -9.5, label = nce_tmie), color="gray", size = 8) 

  theme(strip.text.x = element_text(size = 8),
      strip.text.y = element_text(size = 8), legend.position="none", text = element_text(size = 8)) +
  labs(color= "Study", shape = "Scale")
figgrowth

#ggsave(plot=figgrowth, filename = "fig2.png", width=8, height =4)

# Figure 2
ggsave(plot=figgrowth, device = 'pdf', filename = "fig2_growth.pdf", units = 'cm', width=6, height =6)

```


```{r}

fwgrowth <- rawgrowth %>%
  filter(system == "Freshwater") %>%
  filter(lrr < 2) ## remove WOS outlier

# need to transform

##########

fw_mod <- lmer(lrr~temp + (1 | id), data=fwgrowth)
summary(fw_mod)
Anova(fw_mod)

shapiro.test(lm(lrr~temp,data = fwgrowth)$residuals)
r.squaredGLMM(fw_mod)

##########


confint(fw_mod, level = 0.95)

newdat1 <- data.frame(temp=seq(min(fwgrowth$temp), max(fwgrowth$temp), length=100))

newdat1[,2] <- predict(fw_mod, newdat1, re.form=NA)
newdat1$system <- "Freshwater"

pred1 <- ggpredict(fw_mod, terms = "temp")
pred1$system <- "Freshwater"


##################################################

m.invert <- rawgrowth %>%
  filter(response_type == "growth" & system == "Marine") %>%
  filter(specific_response != "growth efficiency") ## remove this 

str(m.invert)

m.invert.mod <- lmer(lrr~temp + (1 | id), data = m.invert)
summary(m.invert.mod)
Anova(m.invert.mod)

r.squaredGLMM(m.invert.mod)


# not significant

shapiro.test(lm(lrr~temp,data = m.invert)$residuals) 
### failed



# bottomright 


terr <- rawgrowth %>%
  filter(response_type == "growth" & system == "Terrestrial") 
str(terr)

# can we transform this to make it better?

terr$trans <- (terr$lrr - min(terr$lrr))^2
## transforming doesn't seem to be working

terr.mod <- lm(lrr~temp, data = terr)
summary(terr.mod)
Anova(terr.mod)

shapiro.test(terr.mod$residuals) # fail



newdat2 <- data.frame(temp=seq(min(terr$temp), max(terr$temp), length=100))

newdat2[,2] <- predict(terr.mod, newdat2,  re.form=NA)
newdat2$system <- "Terrestrial"

pred2 <- ggpredict(terr.mod, terms = "temp")
pred2$system <- "Terrestrial"


# color palette
N=length(unique(rawgrowth$id))
Cols = colorRampPalette(brewer.pal(8, "Accent"))

# Plot

figgrowth <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=rawgrowth, mapping=aes(x=temp, y=lrr, color=id), 
             position=position_dodge(width=0.8)) +
  #scale_color_viridis_d(option='viridis') +
  scale_color_manual(values=Cols(N))+
  scale_shape_manual(values = c(1,19)) +
  geom_ribbon(pred1, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
   geom_ribbon(pred2, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
   geom_line(data=newdat1, mapping=aes(x=temp, y=V2), color="black") +
  geom_line(data=newdat2, mapping=aes(x=temp, y=V2), color="black") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
    coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(data=rawgrowth, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-sqrt(v), ymax = lrr+sqrt(v)), 
                position=position_dodge(width=0.8)) +
  theme_bw() +
  facet_wrap(~system,nrow=3,ncol=1) +
  theme(strip.text.x = element_text(size = 8),
      strip.text.y = element_text(size = 8), legend.position="none") +
  labs(color= "Study")
figgrowth

ggsave(plot=figgrowth, device = 'pdf', filename = "fig3_growth_system.pdf", units = 'cm', width=6, height =17)

```

## - figure 3a
```{r}


fwgrowthfig <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=fwgrowth, mapping=aes(x=temp, y=lrr, color=id),
             position=position_dodge(width=0.8), size=2) +
  geom_ribbon(pred1, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
   geom_line(data=newdat1, mapping=aes(x=temp, y=V2), color="black") +
  geom_text(data = fwgrowth, aes(x=16, y= -3.9, label = nce_tmie), color="gray", size = 8) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-5,17)+
    coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(data=fwgrowth, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
                position=position_dodge(width=0.8), width = 0.5) +
  theme_bw() +
  facet_wrap(~system) +
  theme(strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale") +
   annotate("text", x = -11, y = 2, label = "A)")
fwgrowthfig


```


## - figure 3b
```{r}
mgrowthfig <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=m.invert, mapping=aes(x=temp, y=lrr, color=id),
             position=position_dodge(width=0.8), size=2) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
     xlim(-5,17)+
 #   coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(data=m.invert, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
                position=position_dodge(width=0.8), width = 1) +
  geom_text(data = m.invert, aes(x=16, y= -3.9, label = nce_tmie), color="gray", size = 8) +
  theme_bw() +
  facet_wrap(~system) +
  theme(strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale") +
   annotate("text", x = -11, y = 2, label = "B)")
            
mgrowthfig


```

## - Figure 3
```{r}

labels <- data.frame(
  system = c("Marine", "Freshwater"),  # your actual system names
  label  = c("B)", "A)")
)

growthfig <- rawgrowth %>%
  filter(system != "Terrestrial")%>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id),
             position=position_dodge(width=0.8), size=2) +
  geom_ribbon(pred1, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
   geom_line(data=newdat1, mapping=aes(x=temp, y=V2), color="black") +
  geom_text(aes(x=10, y= -3.9, label = nce_tmie), color="gray", size = 8) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
   xlim(-5,11)+
    coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
                position=position_dodge(width=0.8), width = 0.5) +
  theme_bw() +
  geom_text(data = labels, aes(x = -4.5, y = 1.9, label = label), size = 5, inherit.aes = FALSE) +
  facet_wrap(~system) +
  theme(strip.text.x = element_text(size = 12),
      strip.text.y = element_text(size = 12), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale") 
growthfig


ggsave(plot=growthfig, filename = "fig3.png", width=10, height =4)

```

# density

```{r}


rawdens <- w_error %>%
  filter(response_type == "density") %>%
  filter(nce_tmie == "TMIE") %>%
  mutate(lrr = if_else(id == "addtl3-2022", -(lrr), lrr)) %>%
  mutate(lrr = if_else(id == "2017D20", -(lrr), lrr)) 


dens_mod <- lmer(lrr~temp + (1 | id), data=rawdens)
summary(dens_mod)
Anova(dens_mod)

r.squaredGLMM(dens_mod)
shapiro.test(lm(lrr~temp,data = rawdens)$residuals)

# color palette
N=length(unique(rawdens$id))
Cols = colorRampPalette(brewer.pal(8, "Accent"))

densfig <- rawdens %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), 
             position=position_dodge(width=0.8), size=2)+
 # scale_color_viridis_d(option='viridis') +
  scale_color_manual(values=Cols(N))+
 # geom_text(aes(x=16, y= -2, label = nce_tmie), color="grey98", size = 8) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Density") +
   xlim(-5,17)+
    coord_cartesian(ylim = c(-2,2)) +  
  geom_errorbar(data=rawdens, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-sqrt(v), ymax = lrr+sqrt(v)), 
                position=position_dodge(width=0.8), width = 0) +
  #  geom_text(data = rawdens, aes(x=16.5, y= -3.9, label = nce_tmie), color="grey", size = 8) +
  theme_bw() +
  theme(strip.text.x = element_text(size = 8),
      strip.text.y = element_text(size = 8), legend.position="none", text = element_text(size = 8)) +
  labs(color= "Study", shape = "Scale")
densfig

ggsave(plot=densfig, filename = "fig4.png", width=8, height =4)

ggsave(plot=densfig, device = 'pdf', filename = "fig4_density.pdf", units = 'cm', width=6, height =6)
# 

```

## - figure 4

```{r}


tdens <- rawdens %>%
  filter(system == "Terrestrial") 

t_mod <- lmer(lrr~temp + (1 | id), data=tdens)
summary(t_mod)
Anova(t_mod)

r.squaredGLMM(t_mod)
shapiro.test(lm(lrr~temp,data = tdens)$residuals)

confint(t_mod, level = 0.95)

newdat3 <- data.frame(temp=seq(min(tdens$temp), max(tdens$temp), length=100))

newdat3[,2] <- predict(t_mod, newdat3, re.form=NA)
newdat3$system <- "Terrestrial"

pred3 <- ggpredict(t_mod, terms = "temp")
pred3$system <- "Terrestrial"
```


## - figure 5

```{r}

# color palette
N=length(unique(tdens$id))
Cols = colorRampPalette(brewer.pal(8, "Accent"))

tdensfig <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=tdens, mapping=aes(x=temp, y=lrr, color=id), 
             position=position_dodge(width=0.8), size=2)+
  geom_ribbon(pred3, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
   geom_line(data=newdat3, mapping=aes(x=temp, y=V2), color="black") +
 # scale_color_viridis_d(option='viridis') +
  scale_color_manual(values=Cols(N))+
 # geom_text(data = tdens, aes(x=15, y= -0.8, label = nce_tmie), color="grey", size = 8) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Density") +
   xlim(-5,17)+
    coord_cartesian(ylim = c(-1,2)) +  
  geom_errorbar(data=tdens, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-sqrt(v), ymax = lrr+sqrt(v)), 
                position=position_dodge(width=0.8), width = 0) +
  theme_bw() +
  facet_wrap(~system) +
  theme(strip.text.x = element_text(size = 8),
      strip.text.y = element_text(size = 8), legend.position="none", text = element_text(size = 10)) +
  labs(color= "Study", shape = "Scale")
tdensfig

 #ggsave(plot=tdensfig, filename = "fig5.png", width=5, height = 4)
 
 ggsave(plot=tdensfig, device = 'pdf', filename = "fig5_density_terrestrial.pdf", units = 'cm', width=6, height =6)


 

```

```{r}


tdens2 <- tdens %>%
  filter(temp < 16)

t_mod <- lmer(lrr~temp + (1 | id), data=tdens2)
summary(t_mod)
Anova(t_mod)

r.squaredGLMM(t_mod)
shapiro.test(lm(lrr~temp,data = tdens2)$residuals)
## ALMOST passes shapiro

confint(t_mod, level = 0.95)

newdat3 <- data.frame(temp=seq(min(tdens$temp), max(tdens$temp), length=100))

newdat3[,2] <- predict(t_mod, newdat3, re.form=NA)
newdat3$system <- "Terrestrial"

pred3 <- ggpredict(t_mod, terms = "temp")
pred3$system <- "Terrestrial"



```







## - supplemental figs/ all other figs together


```{r}

# display negative for density


## amer and vermiert are actually displayed as survival, so these need to be switched too



supp <- w_error %>%
  mutate(lrr = if_else(id == "addtl3-2022", -(lrr), lrr)) %>%
  mutate(lrr = if_else(id == "2017D20", -(lrr), lrr)) %>%
   mutate(lrr = if_else(id == "new6a", -(lrr), lrr)) %>%
  mutate(lrr = if_else(id == "new8a", -(lrr), lrr)) 

## amer and vermiert are actually displayed as survival, so these need to be switched too



supp$response_type <- fct_recode(supp$response_type, "Density" = "density", 
                                        "Fecundity" = "fecundity", "Growth" = "growth", 
                                        "Mortality" = "mortality", 
                                        "Population Growth" = "population growth")

supp <- supp %>%
  mutate(response_facet = paste(response_type, nce_tmie, sep = "_"))

supp$response_facet=factor(supp$response_facet, levels=c("Growth_NCE", "Mortality_NCE",
                                                                     "Fecundity_NCE", 
                                                                     "Population Growth_NCE",
                                                               "Density_NCE", "Density_TMIE"))

facet_labels <- supp %>%
  distinct(response_facet, response_type) %>%
  deframe()
print(facet_labels)
facet_labeller <- as_labeller(facet_labels)

str(supp)

# color palette
N=length(unique(supp$id))
Cols = colorRampPalette(brewer.pal(8, "Accent"))

suppfig <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=supp, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8)) +
  scale_shape_manual(values = c(1,19)) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
     coord_cartesian(ylim = c(-4.1,2)) + 
    xlim(-11,18)+
  geom_errorbar(data=supp, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-sqrt(v), ymax = lrr+sqrt(v)), 
                position=position_dodge(width=0.8), width = 1) +
  geom_text(data = supp, aes(x=11, y= -3.5, label = nce_tmie), color="gray", size = 5) +
  theme_bw() +
 #  scale_color_viridis_d(option='viridis') +
  scale_color_manual(values=Cols(N))+
facet_grid(response_facet ~ system, labeller = labeller(response_facet = as_labeller(
             c("Growth_NCE" = "Growth",
               "Mortality_NCE" = "Mortality",
               "Fecundity_NCE" = "Fecundity", 
             "Population Growth_NCE" = "Populatiom Growth",
               "Density_NCE" = "Density",
               "Density_TMIE" = "Density") 
           ))) +
  theme(strip.text.x = element_text(size = 8),
      strip.text.y = element_text(size = 8), legend.position="none") +
  labs(color= "Study", shape = "Scale")
suppfig


#ggsave(plot=suppfig, filename = "supplement.png", width=5, height = 9)

ggsave(plot=suppfig, device = 'pdf', filename = "figS1_multiple.pdf", units = 'cm', width=12, height =24)



```



## ---------------- cut after here

# trrestrial growth
```{r}


tgrowthfig <- ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=terr, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8)) +
  scale_shape_manual(values = c(1,19)) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,18)+
    coord_cartesian(ylim = c(-4.1,2)) +  
  geom_errorbar(data=terr, mapping=aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
                position=position_dodge(width=0.8), width = 1) +
  theme_bw() +
  facet_wrap(~system) +
  theme(strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10), legend.position="none") +
  labs(color= "Study", shape = "Scale")
tgrowthfig


ggsave(plot=tgrowthfig, filename = "supp.png", width=9, height = 3)

```


#  r population growth

```{r}
rawrep <- w_error%>%
  filter(response_type == "population growth")

# what if we look at this together?

fw.pop <- rawrep %>%
  filter(system == "Freshwater")
str(fw.pop)



quad_mod <- lmer(lrr~temp + I(temp^2) + (1 | id), data=fw.pop)
summary(quad_mod)
Anova(quad_mod)

r.squaredGLMM(quad_mod)


newdat3 <- data.frame(temp=seq(min(fw.pop$temp), max(fw.pop$temp), length=100))

newdat3[,2] <- predict(quad_mod, newdat3,  re.form=NA)
newdat3$system <- "Freshwater"

pred3 <- ggpredict(quad_mod, terms = "temp")
pred3$system <- "Freshwater"

shapiro.test(lm(lrr~temp, data = fw.pop)$residuals) # fail
plot(lm(lrr~temp, data = fw.pop))



# terrestrial

t.pop <- rawrep %>%
  filter(system == "Terrestrial")


terr.mod <- lm(lrr~temp, data = t.pop)
summary(terr.mod)
Anova(terr.mod)

# almost significant

figpop <- ggplot() +
    geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=rawrep, mapping=aes(x=temp, y=lrr, color=id), 
             position=position_dodge(width=0.8)) +
  geom_blank(data=blank, mapping = aes(x=temp, y=lrr)) +
     geom_ribbon(pred3, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_line(data=newdat3, mapping=aes(x=temp, y=V2)) +
  geom_errorbar(data=rawrep, aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
             position=position_dodge(width=0.8), width = 1) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Pop. Growth Rate") +
  theme_bw() +
  xlim(-11,18)+
  coord_cartesian(ylim = c(-0.5, 1.5)) +  
 facet_grid(~system)+
   theme(legend.position="none", strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10)) +
  labs(color= "Study", shape = "Scale")
figpop

ggsave(plot=figpop,filename = "figpop.png", width=9, height = 3)



```


# mortality

```{r}

fw.mort <- w_error %>%
  filter(response_type == "mortality", system == "Freshwater")
str(fw.mort)

fw.mort.mod <- lm(lrr~temp, data = fw.mort)
summary(fw.mort.mod)
Anova(fw.mort.mod)

###############

rawmort <- w_error%>%
  filter(response_type == "mortality") 

rawmort.mod <- lm(lrr~temp, data = rawmort, system == "Terrestrial")
summary(rawmort.mod)
Anova(rawmort.mod)



figmort <- ggplot() +
     geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=fw.mort, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8))+
    geom_errorbar(data=fw.mort, aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v),
             position=position_dodge(width=0.8), width = 1) +
  geom_point(data=rawmort, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8)) +
  geom_errorbar(data=rawmort, aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v),
             position=position_dodge(width=0.8), width = 1) +
    geom_blank(data=blank, mapping = aes(x=temp, y=lrr)) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Mortality") +
  theme_bw() +
    xlim(-11,18)+
 facet_grid(~system)  +
  theme(legend.position="none", strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10)) +
  labs(color= "Study", shape = "Scale")

figmort
ggsave(plot=figmort,filename = "figmort.png", width=9, height = 3)




```

# fecundity 

```{r}

rawfecun <- w_error%>%
  filter(response_type == "fecundity") 


fecun.mod <- lm(lrr~temp, data = rawfecun)
summary(fecun.mod)
Anova(fecun.mod)


shapiro.test(lm(lrr~temp, data = rawfecun)$residuals)  # pass


figfecun <- ggplot() +
     geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=rawfecun, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8)) +
  geom_errorbar(data=rawfecun, aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v), 
             position=position_dodge(width=0.8), width = 1) +
    geom_blank(data=blank, mapping = aes(x=temp, y=lrr)) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Fecundity") +
  theme_bw() +
    xlim(-11,18)+
 facet_grid(~system) +
  theme(legend.position="none", strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10)) +
  labs(color= "Study", shape = "Scale")
figfecun

ggsave(plot=figfecun,filename = "figfecun.png", width=9, height = 3)


```

```{r}


# check individual papers



w_error %>% 
  filter(id == "addtl11-2018") %>%
ggplot() +
     geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id, shape=scale))  


```



# density

```{r}
rawdens <- w_error%>%
  filter(response_type == "density") %>%
  mutate(lrr = if_else(id == "addtl3-2022", -(lrr), lrr)) %>%
  mutate(lrr = if_else(id == "2017D20", -(lrr), lrr)) 

## ok cool, now what if we look at these together? 
## need to switch signs of morton, units are actually algae eaten, addtl3-2022
## need to switch signs of beleznai, which is in plant damage, 2017D20




fw.dens <- rawdens %>%
  filter(system == "Freshwater")
str(fw.dens)


dens.mod <- lm(lrr~temp, data = fw.dens)
summary(dens.mod)
Anova(dens.mod)

# significant

shapiro.test(lm(lrr~temp, data = fw.dens)$residuals) ## almost passes
plot(lm(lrr~temp,  data = fw.dens))

# #almost
# 
newdat4 <- data.frame(temp=seq(min(fw.dens$temp), max(fw.dens$temp), length=100))

newdat4[,2] <- predict(dens.mod, newdat4,  re.form=NA)
newdat4$system <- "Freshwater"


pred4 <- ggpredict(dens.mod, terms = "temp")
pred4$system <- "Freshwater"





t.dens <- rawdens %>%
  filter(system == "Terrestrial") # not significant even without bannerman
str(t.dens)

dens.mod2 <- lmer(lrr~temp+ (1 | id),  data = t.dens)
summary(dens.mod2)
Anova(dens.mod2)

r.squaredGLMM(dens.mod2)


# almost significant

newdat5 <- data.frame(temp=seq(min(t.dens$temp), max(t.dens$temp), length=100))

newdat5[,2] <- predict(dens.mod2, newdat5,  re.form=NA)
newdat5$system <- "Terrestrial"


pred5 <- ggpredict(dens.mod2, terms = "temp")
pred5$system <- "Terrestrial"



figdens <- ggplot() +
     geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=rawdens, mapping=aes(x=temp, y=lrr, color=id, shape=scale), 
             position=position_dodge(width=0.8)) +
  geom_errorbar(data=rawdens, aes(x=temp, y=lrr, color=id, ymin = lrr-v, ymax = lrr+v),
              position=position_dodge(width=0.8), width = 1) +
  geom_ribbon(pred4, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_ribbon(pred5, mapping=aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) +
    geom_line(data=newdat4, mapping=aes(x=temp, y=V2), color="black") +
    geom_line(data=newdat5, mapping=aes(x=temp, y=V2), color="grey") +
    geom_blank(data=blank, mapping = aes(x=temp, y=lrr)) +
  #  geom_line(data=newdat4, mapping=aes(x=temp, y=V2)) +
  scale_shape_manual(values = c(1,19)) +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Density") +
  theme_bw() +
    xlim(-11,18)+
 facet_grid(~system) +
  theme(legend.position="none", strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10)) +
  labs(color= "Study", shape = "Scale")

figdens

ggsave(plot=figdens,filename = "figdens.png", width=9, height = 3)



```

###### testing from DK


```{r}

stoks <- rawgrowth %>% 
  filter(id == "2012Ex1-69") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Stoks") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
stoks

culler <- rawgrowth %>% 
  filter(id == "2014D9") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("culler") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
culler

Janssens <- rawgrowth %>% 
  filter(id == "2015D4") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Janssens") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Janssens

CorreaAraneda <- rawgrowth %>% 
  filter(id == "2017E38") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Correa-Araneda") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
CorreaAraneda


Burraco <- rawgrowth %>% 
  filter(id == "addtl10-2018") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Burraco") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Burraco



Amer <- rawgrowth %>% 
  filter(id == "new6") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Amer") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Amer



Vermiert <- rawgrowth %>% 
  filter(id == "new8") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Vermiert") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Vermiert


Stahlschmidt <- rawgrowth %>% 
  filter(id == "new9") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Stahlschmidt") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Stahlschmidt

Wos <- rawgrowth %>% 
  filter(id == "new10") %>%
  ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(mapping=aes(x=temp, y=lrr, color=id), size = 5) +
  ggtitle("Wos") +
  xlab("Temperature Increase") +
  ylab("LRR of Predation Risk on Growth") +
    xlim(-11,11)+
  theme_bw() +
  theme(strip.text.x = element_text(size = 15),
      strip.text.y = element_text(size = 15), legend.position="none", text = element_text(size = 15)) +
  labs(color= "Study", shape = "Scale")
Wos

```


```{r}


ggplot() +
  geom_hline(yintercept=0, color="darkgrey") +
  geom_point(data=supp, mapping=aes(x=temp, y=lrr, color=id, shape=scale)) +
facet_wrap(facets="id")


```


